Lec_2_4_2014
========================================================
Warm-up for today's class
From this article:
http://www.nature.com/mp/journal/vaop/ncurrent/full/mp2013178a.html

Pros
- small multiples (similar structure) : easy to interprete
- 

Cons
- scale differences
- bar plot? (much more ink), how about using boxplot as an alternitive?, skewed distru
- selected p-values (p=0.072 in h)

Final concepts of SLR

- R^2 = 1 - proportional of all variance described by erros
      = proportion of variance explained by model
      = squre of correlation (only for simple linear regression) - recheck !
      = NEVER used for hypothesis test
      = 1 - SS(res)/SS(tot) = SSreg/SStot

- ANOVA = SStot == SS(reg) + SS(res) (See textbook)

- R command : names(), class(), resid(), fitted(), summary(object)$coef, anova()

- Centering => make intercept be meaningful, no effect on slope!

- rescaling => change in slope but not changes in intepretation (inference won't change)

- 

```{r}

```



```{r fig.width=7, fig.height=6}

```

