Problem Set_2
3.7.2014
Sangsoo Park 
========================================================

•  Create a short reproducible document (using knitr) that describes the basic structure of a dataset and summarizes some key features of the data using a few key tables and figures. Choose a dataset from these datasets or the ones in the class Google Drive. If your dataset has a lot of variables, focus on a subset of them – less than 6 or so – for the purposes of this exercise. Your write-up should answer the following questions:

o	What is the background/context for this data?

    - To understand factors that result in diabetes of African American in Central Virginia. 
    - To figure out what variables are related to the prevalence of diabetes of African American in Central Virginia. 

o	How many observations are there?

    - 19 variables from 403 subjects who have diabetes
    - Data were collected by interview

o	What is the unit of observation?

    - Height in inch, Weight in lb, length of waist and hip in inch, age in year
    - No ideas of units for the other variables such as hdl (mmol/L or mg/L), glyhd, etc 

o	Is there any missing data? If so, are there patterns to the missingness?

    - For weight variable, there was one missing data 
    - For hip variable, there was two missing data
    - For glyhb variable, there was 13 missing data
    - I don't think that the missing data has pattern.  

```{r}
# Load library
library(Hmisc)

# Load Diabete dateset
getHdata(diabetes)

sub_diabetes <- subset(diabetes, select = c(chol,glyhb,weight,gender,waist,hip))
which(is.na(sub_diabetes$glyhb))
```

o	What are the key variables and what do their distributions look like?

    - Waist/ hip ratio : The study found that the ratio might be related to prevalence of diabetes. 
    - Length of waist and hip : we define obesity using the information
    - Distributions of waist and hip look like the normal distribution
  
```{r}
hist(sub_diabetes$waist)
hist(sub_diabetes$hip)
```

o	Is there a pair of variables that might work well for a Simple Linear Regression? (You don't necessarily need to run one, but you could.)

    - The slope is so small between weight and glyhb.
    - The slope is still small but relatively larger between hip and glyhb.
```{r}
model1 <- lm(glyhb ~ hip, data=sub_diabetes)
summary(model1)
plot(sub_diabetes$hip, sub_diabetes$glyhb)
abline(model1, col="red")
```

o	Are there any obvious outliers in the data?

    - For the Weight variable, the mean was around 175 lb and the standard deviation was around 38 lb. There was an outlier, over 290 lb, which is larger than the value from summation of the mean and 3 * standard deviation. 
```{r}
# Boxplot to check outliers of the "weight" variable
boxplot(sub_diabetes$weight, xlab="Weight", ylab="lb")
```

o Add one or two simple regression 
```{r}
# between chol and glyhb
model2 <- lm(glyhb ~ chol, data=sub_diabetes)
summary(model2)
plot(sub_diabetes$chol, sub_diabetes$glyhb)
abline(model2, col="red")

# between weight and glyhb
model <- lm(glyhb ~ weight, data=sub_diabetes)
summary(model)
plot(sub_diabetes$weight, sub_diabetes$glyhb)
abline(model, col="red")
```

o Creat "slr" function

```{r}
# Testing slr function to compare lm function and hand calculation of b0 and b1
slr <- function(x,y){
  
  # Using lm function
  r_model <- lm(y ~ x)
  
  # Before calculation, remove rows including missing values
  raw_data <- data.frame(x,y)
  pro_data <-  na.omit(raw_data)
  
  x <- pro_data[,1]
  y <- pro_data[,2]
  # Using calculation from textbook p18
  res_x <- x - mean(x)  
  res_y <- y - mean(y)
  res_xy <- sum(res_x * res_y)
  squared_res_x <- res_x^2
  
  b1_hand <- sum(res_xy)/sum(squared_res_x)
  b0_hand <- mean(y) - b1_hand*mean(x)
  
  hand_coef <- c(b0_hand,b1_hand)
  names(hand_coef) <- c("Intercept","x")
  
  # Return results to compare between the two methods
  result <- list(r_model$coef,hand_coef)
  names(result)[1] <- "lm_function"
  names(result)[2] <- "hand_cal"
  return(result)
}
```

o Comparison, 
  
```{r}
# See that the function gives us the exact same result
slr(sub_diabetes$weight, sub_diabetes$glyhb)
```

o MLR, equation, interpretation of coefficient

```{r}
# Add mulitiple regression
multi_prac <- lm(formula =  glyhb~ weight + waist + hip, data = sub_diabetes)
summary(multi_prac)
```

  - Equation
$$y_i = \beta_0 + \beta_1x_{i1} + \beta_2x_{i2} + \beta_3x_{i3} + \epsilon_i; \mbox{ } \epsilon_i \stackrel{iid}{\sim} (0,\sigma^2)$$

  - Based on the result from MLR,
$$y_i = 2.41 - 0.005*x_{i1} + 0.16*x_{i2} - 0.05*x_{i3}$$
  
  - Interpretation
  
  A unit increase in waist increased 0.16 of expected glyhb while the other variables are constant. We can interprete the other variables by the same way.
  $\beta_0$ indicates expected value of y when all x variables are zero.
      
o MLR including a categorical variable      
```{r}
multi_cate <- lm(formula =  glyhb~ chol + hip + gender, data = sub_diabetes)
summary(multi_cate)
```

  - Results
  
  When female (gender = 1)
  $$glyhb_i = 0.93 + 0.01*chol_{i1} - 0.06*hip_{i2} - 0.45$$
  $$glyhb_i = 0.48 + 0.01*chol_{i1} - 0.06*hip_{i2} $$
  When male (gender = 0)
  $$glyhb_i = 0.93 + 0.01*chol_{i1} - 0.06*hip_{i2}$$
  
  - Interpretation
  $\beta_0 indicates expected value of glyhb when all x variables are zero. When gender is female, the intercept of MLR is decreased by 0.45. When gender is male, the MLR is explained by two variables, chol and hip because x value for gender is zero. 